Containers and R---the Rockerverse and beyond

### 2020-01-18 manuscript.Rmd
sneaking into the list of authors
fix formatting

One of the first implementations of having R respond to client reqests were TCP/IP and local socket servers, such as \pkg{Rserve} ([https://cran.r-project.org/package=Rserve](https://cran.r-project.org/package=Rserve)) and \pkg{svSocket} ([https://cran.r-project.org/package=svSocket](https://cran.r-project.org/package=svSocket)), which allowed other local or remote processes, even written in other programming languages, such as C/C++ or Java, to make binary requests to interact with R without the need to link to R at compile time.

R could also empower webserver since at least 2007, when [RApache](http://www.rapache.net) was presented at the useR! conference. Rapache runs an R interpreter inside the Apache web server and makes it possible and very efficient to build scalable R-based APIs and web services.

A more recent related work is the \pkg{plumber} ([https://www.rplumber.io/](https://www.rplumber.io/)) package, which makes it even easier for R to be used to host web services through RESTful APIs.
When R is running on a server, a different server can ask R to make a computation through an HTTP request and get an immediate response.
For instance, if a data scientist at a company creates a machine learning model to tell which customers are likely to make a repeat purchase, if they serve the model through a plumber API then someone else on the marketing team could write software that sends different emails depending on that real-time prediction.
Plumber syntax is simple to read and understand: the API is created by commenting above function names in an R script.

@@

\label{azurecontainers}
**\pkg{AzureContainers}** is an umbrealla package providing interfaces the three services of Microsoft's Azure Cloud, namely [Container Instances](https://azure.microsoft.com/en-us/services/container-instances/) for running individual containers, [Container Registry](https://azure.microsoft.com/en-us/services/container-registry/) for private image distribution, and [Kubernetes Service](https://azure.microsoft.com/en-us/services/kubernetes-service/) for orchestrated deployments.
While a package like \pkg{plumber} provides the infrastructure for turning an R workflow a service, for production purposes it is usually necessary to take into account scalability, reliability and ease of management.
A single container is limited in the volume of requests it can service, and if the container goes down or is occupied with processing a computationally intensive task, the service becomes unavailable.
These issues can be mitigated by deploying not a single container, but a _cluster_ of containers, orchestrated as a single deployment, e.g., with [Kubernetes](https://kubernetes.io).
A cluster features load-balancing, autoscaling of containers across numerous servers (in the cloud or on premise), and restarting failed ones.
\pkg{AzureContainers} provides a lightweight yet powerful interface for working with Resource Manager, the framework for deploying and managing arbitrary resources in Azure.
On the client side, it provides simple shells to the Docker, and to Kubernetes' `kubectl` and `helm` tools, that are needed for managing Kubernetes clusters.
The package simplifies complex infrastructure management to a number of R funcction calls, given an Azure account with sufficient credit\footnots{See _"Deploying a prediction service with Plumber"_ vignette at \href{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01_plumber_deploy.html}{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01_plumber_deploy.html} for details.}.

>>>>

The backbone of this work is the \pkg{plumber} ([https://www.rplumber.io/](https://www.rplumber.io/)) package.
The plumber package allows R to be used to host web services through RESTful APIs.
When R is running on a server, a different server can ask R to make a computation through an HTTP request and get an immediate response.
For instance, if a data scientist at a company creates a machine learning model to tell which customers are likely to make a repeat purchase, if they serve the model through a plumber API then someone else on the marketing team could write software that sends different emails depending on that real-time prediction.
Plumber syntax is simple to read and understand: the API is created by commenting above function names in an R script.

@@

\label{azurecontainers}
**\pkg{AzureContainers}** is an umbrealla package providing interfaces the three services of Microsoft's Azure Cloud, namely [Container Instances](https://azure.microsoft.com/en-us/services/container-instances/) for running individual containers, [Container Registry](https://azure.microsoft.com/en-us/services/container-registry/) for private image distribution, and [Kubernetes Service](https://azure.microsoft.com/en-us/services/kubernetes-service/) for orchestrated deployments.
While a package like \pkg{plumber} provides the infrastructure for turning an R workflow a service, for production purposes it is usually necessary to take into account scalability, reliability and ease of management.
A single container is limited in the volume of requests it can service, and if the container goes down or is occupied with processing a computationally intensive task, the service becomes unavailable.
These issues can be mitigated by deploying not a single container, but a _cluster_ of containers, orchestrated as a single deployment, e.g., with [Kubernetes](https://kubernetes.io).
A cluster features load-balancing, autoscaling of containers across numerous servers (in the cloud or on premise), and restarting failed ones.
\pkg{AzureContainers} provides a lightweight yet powerful interface for working with Resource Manager, the framework for deploying and managing arbitrary resources in Azure.
On the client side, it provides simple shells to the Docker, and to Kubernetes' `kubectl` and `helm` tools, that are needed for managing Kubernetes clusters.
The package simplifies complex infrastructure management to a number of R funcction calls, given an Azure account with sufficient credit\footnote{See \emph{"Deploying a prediction service with Plumber"} vignette for details:  \href{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01_plumber_deploy.html}{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01\_plumber\_deploy.html}.}.
